---
title: "Final project"
author: "Daniel Gonzalez-Suarez"
date: "4/3/2020"
output: 
  pdf_document: 
    toc: yes
  html_document: 
    theme: united
    toc: yes
editor options:
  chunk_output_style: console
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Part II, Final Assignment
This is the second assingment for the Ecological Data Analysis in R course. In order to start, packages and functions will be loaded first.
```{r, include=FALSE}
library(tidyverse)
library(dplyr)
library(vegan)
source("./functions/6003Functions.R")
```

## Loading data and tidying it up.

Here, steps for tidying the dataset will be presented. Raw data was taken during completion of underwater transects, then input in Excel. Data will be imported as a .csv file.  

```{r}
df <- read.csv("./data/df_daniel.csv")

#df1 <- summarise(df, Provincia) 
#ab_matrix <- pivot_wider(df1, names_from = c("Especie", "EspecieCodigo"), values_from = "Abundancia")

# Create spp abundance matrix
# this matrix only needs three data: spp names, abundance values and site/location/province to allocate specific area sampled
# therefore we will first select these variables
temp_data <- df %>% select(Provincia, Localidad, SitioCodigo, Especie, Abundancia)
head(temp_data)
str(temp_data)

# let´s now summarise our spp regardless of their size (variable which we ignored in the previous step)
temp_data <- summarise(group_by(temp_data, Provincia, Localidad, SitioCodigo, Especie),
                        sum_abundancia = sum(Abundancia, na.rm = TRUE))

# let´s now make the abundance matrix, we indicate variable names and abundance values
# rest of variables will be "ignored" but maintained
ab_matrix <- pivot_wider(temp_data, names_from = Especie, values_from = sum_abundancia)

# The lines of code above work fine, my only observation like I mentioned during our meeting is
# that you only have seven data points, so for a linear model may be too few
# but if it is too hard and time consuming to go back to a longer dataset version, 
# you can run the excercise like this for your final project =)
# let me know if you have any other questions

# I did not revise any of the following lines of code
head(df)
tail(df)
str(df)
range(df$Talla_.cm.)
range(df$Abundancia)
range(df$Densidad_.org.m2.)
```

Everything looks good. Data was already in long format before importing it. 


## Exploring the data
We are going to see visual representation of data.

# 1. Outliers 

```{r}
boxplot(df$Talla_.cm.)
boxplot(df$Abundancia)
boxplot(df$Densidad_.org.m2.)

dotchart(df$Talla_.cm.)
dotchart(df$Abundancia)
dotchart(df$Densidad_.org.m2.)
```

Many outliers, but there are no errors on data.

# 2. Homogeneity of variance 

```{r}
p <- ggplot(data=df, aes(X=Provincia, y=Talla_.cm.)) +
  geom_boxplot() + facet_wrap(~SitioCodigo)
p
p<- ggplot(data=df, aes(X=SitioCodigo, y=Densidad_.org.m2.)) +
  geom_boxplot() + facet_wrap(~Localidad)
p
```

There is homogeneity of variance.

# 3. Normality 

```{r}

hist(df$Talla_.cm.)
hist(df$Abundancia)
hist(df$Densidad_.org.m2.)

qqnorm(df$Talla_.cm.)
qqline(df$Talla_.cm.)

qqnorm(df$Abundancia)
qqline(df$Abundancia)

qqnorm(df$Densidad_.org.m2.)
qqline(df$Densidad_.org.m2.)

```

Data is not normal, as seen in histograms and confirmed by Q-Q plots.

# 4. Zeroes 

```{r}
range(df$Talla_.cm.)
plot(table(df$Talla_.cm.))

range(df$Abundancia)
plot(df$Abundancia)

range(df$Densidad_.org.m2.)
plot(table(df$Densidad_.org.m2.))

```

There are no zeroes in either continuous variable.


# 5. Collinearity X
```{r}
pairs(~ Provincia + Localidad + SitioCodigo + Densidad_.org.m2. + Talla_.cm., 
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      data=df) 

plot(Talla_.cm. ~ Provincia, data=df)
plot(Talla_.cm. ~ Localidad, data=df)
plot(Talla_.cm. ~ SitioCodigo, data=df)

plot(Densidad_.org.m2. ~ Provincia, data=df)
plot(Densidad_.org.m2. ~ Localidad, data=df)
plot(Densidad_.org.m2. ~ SitioCodigo, data=df)

pairs(~ Provincia + Densidad_.org.m2. + Talla_.cm., 
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      data=df) 
```

Our X variable will now be only Province

# 6. Relationships X and Y
```{r}
p <- ggplot(df, aes(x=Provincia, y=Talla_.cm.)) +
  geom_boxplot()
p

p <- ggplot(df, aes(x=Provincia, y=Densidad_.org.m2.)) +
  geom_boxplot()
p
```


# 7. Interactions

```{r}

coplot(Talla_.cm. ~ Densidad_.org.m2. | Provincia ,
       data=df,
       panel=function(x,y,...) {
         panel.smooth(x,y,span=0.8,iter=5,...)
         abline(lm(y ~ x), col="blue") } )
```
There seems to be no interaction in Provinces. 

# 8. Independence of Y
Both length and density could be dependent on Province because of fishing pressure effects. 

